Manuscript 1 code

### 2024-10-11 manuscript_1_code.Rmd
created option for fig 5 that includes plot level richness-synchrony relationships

```{r richness synchrony mod}
rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
summary(rich_stab_mod)
Anova(rich_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# richness_mean         2721.438  1  < 2.2e-16 ***
# habitat                325.876  3  < 2.2e-16 ***
# richness_mean:habitat   33.926  3  2.054e-07 ***
hist(residuals(rich_stab_mod)) # looks good
plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
# contrast                     estimate     SE  df z.ratio p.value
# Fringing - Backreef          0.186670 0.0414 Inf   4.512  <.0001
# Fringing - Forereef 10m      0.187196 0.0455 Inf   4.114  0.0002
# Fringing - Forereef 17m      0.027112 0.0463 Inf   0.586  0.9364
# Backreef - Forereef 10m      0.000525 0.0405 Inf   0.013  1.0000
# Backreef - Forereef 17m     -0.159559 0.0411 Inf  -3.885  0.0006
# Forereef 10m - Forereef 17m -0.160084 0.0439 Inf  -3.647  0.0015
```

@@

@@

```{r mean synchrony}
mean_synchrony_mod <- glmmTMB(cover_stability ~ synchrony_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
hist(residuals(mean_synchrony_mod)) # chunky
plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # fine
summary(mean_synchrony_mod)
Anova(mean_synchrony_mod)
#                         Chisq Df Pr(>Chisq)    
# synchrony_mean         32.914  1  9.634e-09 ***
# habitat                18.523  3  0.0003430 ***
# synchrony_mean:habitat 19.062  3  0.0002654 ***
emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            -3.26 1.503 Inf  -2.166  0.1327
# Fringing - Forereef 10m         2.22 0.995 Inf   2.231  0.1149
# Fringing - Forereef 17m         3.37 1.220 Inf   2.764  0.0291
# Backreef - Forereef 10m         5.48 1.622 Inf   3.377  0.0041
# Backreef - Forereef 17m         6.63 1.769 Inf   3.748  0.0010
# Forereef 10m - Forereef 17m     1.15 1.363 Inf   0.846  0.8322
```
```{r plot mean synchrony}
mean_synch_emm <- emmip(mean_synchrony_mod,
                        habitat ~ synchrony_mean,
                        at = list(synchrony_mean = seq(0,1,0.01)), plotit = F, CIs = T)

@@

## panel graph
```{r figure 5 panel, fig.height = 8, fig.width=16}
(figure_5 <- ggarrange(figure_5a, figure_5b, 
                      ncol = 2, nrow = 1, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      labels = c("a.", "b."),
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_5.png", figure_5, height = 8, width = 16)
```

>>>>

```{r richness stability mod}
rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
summary(rich_stab_mod)
Anova(rich_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# richness_mean         2721.438  1  < 2.2e-16 ***
# habitat                325.876  3  < 2.2e-16 ***
# richness_mean:habitat   33.926  3  2.054e-07 ***
hist(residuals(rich_stab_mod)) # looks good
plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
# contrast                     estimate     SE  df z.ratio p.value
# Fringing - Backreef          0.186670 0.0414 Inf   4.512  <.0001
# Fringing - Forereef 10m      0.187196 0.0455 Inf   4.114  0.0002
# Fringing - Forereef 17m      0.027112 0.0463 Inf   0.586  0.9364
# Backreef - Forereef 10m      0.000525 0.0405 Inf   0.013  1.0000
# Backreef - Forereef 17m     -0.159559 0.0411 Inf  -3.885  0.0006
# Forereef 10m - Forereef 17m -0.160084 0.0439 Inf  -3.647  0.0015
```

@@

## diversity-synchrony plot level

```{r ds plot}
rich_synch_mod <- glmmTMB(synchrony_trans~ richness_mean*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
summary(rich_synch_mod)
Anova(rich_synch_mod)
#                         Chisq Df Pr(>Chisq)    
# richness_mean         395.416  1  < 2.2e-16 ***
# habitat               136.859  3  < 2.2e-16 ***
# richness_mean:habitat  90.031  3  < 2.2e-16 ***
hist(residuals(rich_synch_mod)) # looks good
plot(residuals(rich_synch_mod) ~ fitted(rich_synch_mod)) 
r.squaredGLMM(rich_synch_mod) # 0.71, 0.78
emtrends(rich_synch_mod, pairwise ~ habitat, var = "richness_mean") # forereefs not different from each other
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           -0.823 0.161 Inf  -5.113  <.0001
# Fringing - Forereef 10m       -1.443 0.183 Inf  -7.879  <.0001
# Fringing - Forereef 17m       -1.628 0.187 Inf  -8.723  <.0001
# Backreef - Forereef 10m       -0.620 0.151 Inf  -4.103  0.0002
# Backreef - Forereef 17m       -0.805 0.154 Inf  -5.240  <.0001
# Forereef 10m - Forereef 17m   -0.185 0.165 Inf  -1.124  0.6745

rich_synch_emm <- emmip(rich_synch_mod,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_synch_effects <- rich_synch_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)

(rich_synch_plot <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = richness_mean, y = synchrony_trans, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_synch_effects, aes(x = richness_mean, y = inv_logit(yvar), colour = habitat), size = 1.5) +
  geom_ribbon(data = filtered_rich_synch_effects, aes(x = richness_mean, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Species synchrony\n", x = "Taxonomic richness", title = "") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=22),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
```

@@

```{r mean synchrony}
mean_synchrony_mod <- glmmTMB(cover_stability ~ synchrony_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
hist(residuals(mean_synchrony_mod)) # chunky
plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # fine
summary(mean_synchrony_mod)
Anova(mean_synchrony_mod)
#                         Chisq Df Pr(>Chisq)    
# synchrony_mean         32.914  1  9.634e-09 ***
# habitat                18.523  3  0.0003430 ***
# synchrony_mean:habitat 19.062  3  0.0002654 ***
emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            -3.26 1.503 Inf  -2.166  0.1327
# Fringing - Forereef 10m         2.22 0.995 Inf   2.231  0.1149
# Fringing - Forereef 17m         3.37 1.220 Inf   2.764  0.0291
# Backreef - Forereef 10m         5.48 1.622 Inf   3.377  0.0041
# Backreef - Forereef 17m         6.63 1.769 Inf   3.748  0.0010
# Forereef 10m - Forereef 17m     1.15 1.363 Inf   0.846  0.8322
```

```{r plot mean synchrony}
mean_synch_emm <- emmip(mean_synchrony_mod,
                        habitat ~ synchrony_mean,
                        at = list(synchrony_mean = seq(0,1,0.01)), plotit = F, CIs = T)

@@

## panel graph
```{r figure 5 panel, fig.height = 6, fig.width=22}
(figure_5 <- ggarrange(rich_synch_plot, figure_5b, figure_5a, 
                      ncol = 3, nrow = 1, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      labels = c("a.", "b.", "c."),
                      font.label = list(size = 26, color = "black", face = "plain")))

# ggsave(filename = "output/figure_5_option2.png", figure_5, height = 8, width = 16)
```
